Debunking Neural Essay Scoring

### 2018-05-12 paper/acl17-latex/paper.tex
Added some minor changes

Recently due to a lot of surge in deep-learning, a lot of neural models have been proposed which claim to beat the human level performance on evaluation metrics. Unlike traditional systems these systems, accepts an essay text as input directly and learns the features automatically from the data. A lot of models with varying architectures have been developed in the recent years. Some of them use a long short-term memory network (LSTM) \cite{taghipour2016neural}, while others use a convolutional neural networks (CNN) \cite{dong2016automatic} for the effect of automatically learning features. Some of them also use an augmented method to learn word embeddings \cite{alikaniotis2016automatic}.

@@

We tried to analyze the effect of the length of the essay on the prediction of the models. Typically, low scoring essays have low scores and this can be seen from \cref{fig:length_plot}. Let, $\mu_l$ denote the mean length of the essays in the test set and $\sigma_l$ denote the standard deviation of the lengths. We selected essays from each prompt that had length less than $\mu_l - 2\sigma_l$ and appended the same essay till the point where the length of the essay exceeded $\mu_l - 2\sigma_l$. The mean score of these essays before and after increasing length were calculated and the percentage increase in the score was calculated. \cref{length} shows that the average increase in the predicted score of the essay is $100.58\%$ while for some test sets the increase in the score was close to $235.53\%$.

@@

\subsection{Effect of Spelling Mistakes}
We tried to analyze the effect of the spelling mistakes on the prediction of the models. We randomly chose 20\% of the words of the essay and then randomly shuffled the characters of the words to generate words with wrong spellings. We had a set of predicted scores for the essays with correct spellings and a new set of predicted scores for essays with wrong spellings. We calculated QWK between these two sets of scores. \cref{spell_word_sentence} shows that for the neural model, the QWK is close to 1 for most prompts whereas it is very close to 0 for non-neural model. Thus, for the neural model there is no disagreement between the scores predicted for essays with correct spellings and essays with 20\% wrong spellings.

@@

\subsection{Effect of Word Shuffle}
We tried to analyze the effect of the bad grammar on the prediction of the models. We randomly shuffled the words of each sentence in the essay. This resulted in an essay with terrible grammar, and thus should be scored low by any good AES system. We had a set of predicted scores for the essays with good grammar and a new set of predicted scores for essays with bad grammar and a lot of grammatical mistakes. We calculated QWK between these two sets of scores. \cref{spell_word_sentence} shows that for the neural model, the QWK is close to 1 for most prompts whereas it is significantly less than 1 for the non-neural model. Thus, for the neural model there is no disagreement between the scores predicted for essays with correct grammar and essays with bad grammar.

@@

\section{Conclusions}
In this paper we have analyzed the behavior of neural model for automated essay scoring on various practical scenarios. We performed various experiments for effect of length, grammar, spelling, organization and found that neural models perform very poorly in these qualitative metrics even while giving state of art performance on the quantitative benchmarks. This shows that the problem of automated essay scoring is far from being solved and getting human level performance on benchmarks is not the solution. This also shows that the current benchmarks are not sufficient for evaluation and thus call for the need of better metrics. We have also suggested some solutions to problems like length-score inflation, misspelled words, but still questions whether these systems can be deployed in real life.

>>>>

Recently due to surge in deep-learning, a lot of neural models have been proposed which claim to beat the human level performance on evaluation metrics. Unlike traditional systems these systems, accepts an essay text as input directly and learns the features automatically from the data. A lot of models with varying architectures have been developed in the recent years. Some of them use a long short-term memory network (LSTM) \cite{taghipour2016neural}, while others use a convolutional neural networks (CNN) \cite{dong2016automatic} for the effect of automatically learning features. Some of them also use an augmented method to learn word embeddings \cite{alikaniotis2016automatic}.

@@

We tried to analyze the effect of the length of the essay on the prediction of the models. Typically, low scoring essays have low scores and this can be seen from \cref{fig:length_plot}. Let, $\mu_l$ denote the mean length of the essays in the test set and $\sigma_l$ denote the standard deviation of the lengths. We selected essays from each prompt that had length less than $\mu_l - 2\sigma_l$ and appended the same essay till the point where the length of the essay exceeded $\mu_l + 2\sigma_l$. The mean score of these essays before and after increasing length were calculated and the percentage increase in the score was reported. \cref{length} shows that the average increase in the predicted score of the essay is $100.58\%$ while for some test sets the increase in the score was close to $235.53\%$.

@@

\subsection{Effect of Spelling Mistakes}
We tried to analyze the effect of the spelling mistakes on the prediction of the models. We randomly chose 20\% of the words of the essay and then did a random shuffle of the characters to generate words with wrong spellings. We had a set of predicted scores for the essays with correct spellings and a new set of predicted scores for essays with wrong spellings. We calculated QWK between these two sets of scores. \cref{spell_word_sentence} shows that for the neural model, the QWK is close to 1 for most prompts whereas it is very close to 0 for non-neural model. Thus, for the neural model there is no disagreement between the scores predicted for essays with correct spellings and essays with 20\% wrong spellings.

@@

\subsection{Effect of Word Shuffle}
We tried to analyze the effect of the bad grammar on the prediction of the models. We randomly shuffled the positions of words in each sentence of the essay set. This resulted in an essay with terrible grammar, and thus should be scored low by any good AES system. We had a set of predicted scores for the essays with good grammar and a new set of predicted scores for essays with bad grammar and a lot of grammatical mistakes. We calculated QWK between these two sets of scores. \cref{spell_word_sentence} shows that for the neural model, the QWK is close to 1 for most prompts whereas it is significantly less than 1 for the non-neural model. Thus, for the neural model there is no disagreement between the scores predicted for essays with correct grammar and essays with bad grammar.

@@

\section{Conclusions}
In this paper we have analyzed the behavior of neural model for automated essay scoring on various practical scenarios. We performed various experiments for effect of length, grammar, spelling, organization and found that neural models perform very poorly in these qualitative metrics even while giving state of art performance on the quantitative benchmarks. This shows that the problem of automated essay scoring is far from being solved and getting human level performance on benchmarks is not the solution. This also shows that the current benchmarks are not sufficient for evaluation and thus call for the need of better metrics. We have also suggested some solutions to problems like length-score correlation, misspelled words which seem to provide some robustness to above problems.
